HSF Community White Paper on Data Organization, Management and Access (DOMA)

### 2018-03-22 CWP/papers/HSF-CWP-2017-04_doma/latex/hsf-cwp-dm.tex
DOMA CWP: LaTeX cleanup

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \begin{quote}
  \textbf{Big Data:} The HL-LHC era will significantly increase both the
  date rate and the data volume. The computing systems will need to
  handle this without significant cost increases and within evolving
  storage technology limitations.
  \end{quote}
\item
  \begin{quote}
  \textbf{Dynamic Distributed Computing:} In addition, the significantly
  increased computational requirements for the HL-LHC era will also
  place new requirements on data. Specifically, the use of new types of
  compute resources (\emph{e.g.,} cloud, HPC) with different dynamic
  availability and characteristics are used will require more dynamic
  DOMA systems.
  \end{quote}
\item
  \begin{quote}
  \textbf{New Applications:} New applications, such as machine learning
  training or high-rate data-query systems for analysis, will likely be
  employed to meet the computational constraints and to extend the
  physics reach of the HL-LHC. These new applications will place new
  requirements on how and where data is accessed and produced. For
  example, specific applications (e.g. training for machine learning)
  may require use of specialized processor resources such as GPUs,
  placing further requirements on data.
  \end{quote}
\end{enumerate}

@@

\begin{itemize}
\item
  \begin{quote}
  \textbf{Data Organization:} This is essentially how data is structured
  as it is written. Most data is written in flat files, in ROOT format,
  typically with a column-wise organization of the data. The records
  corresponding to these columns are compressed. The internal details of
  this organization are typically visible only to individual software
  applications.
  \end{quote}
\item
  \begin{quote}
  \textbf{Data Management:} The key challenge here was the transition to
  the use of distributed computing in the form of the grid. The
  experiments developed dedicated data transfer and placement systems,
  along with catalogs, to move data between computing centers. To first
  order, the computing models were rather static: data was placed at
  sites and the relevant compute jobs were sent to the right locations.
  Applications might interact with catalogs or, at times, the workflow
  management systems does this on behalf of the applications.
  \end{quote}
\item
  \begin{quote}
  \textbf{Data Access:} Various protocols are used for direct reads
  (rfio, dcap, xrootd, etc.) with a given computer center and/or
  explicit local stage-in and caching for read by jobs. Application
  access may use different protocols than those used by the data
  transfers between sites.
  \end{quote}
\end{itemize}

@@

\item
  \begin{quote}
  \textbf{Reconstruction:} Event reconstruction is CPU limited, due to

@@

  \end{quote}
\item
  \begin{quote}
  \textbf{Simulation:} HEP event simulation strategies focus on

@@

  \end{quote}
\item
  \begin{quote}
  \textbf{Analysis:} Analysis level datasets and subsets are typically

@@

  \end{quote}
\item
  \begin{quote}
  \textbf{Replication:} Data replication, whether within a site or

@@

  \end{quote}

@@

\begin{itemize}
\item
  \begin{quote}
  \protect\hypertarget{_xbbh0qorjpnn}{}{}Quantify impact on performances
  and resource utilization (storage, network) for the main type of
  access patterns (simulation, reconstruction, analysis)
  \end{quote}
\end{itemize}

\begin{itemize}
\item
  \begin{quote}
  \protect\hypertarget{_ox25y1bzr703}{}{}Assess impact on catalogs and
  data distribution
  \end{quote}
\end{itemize}

\begin{itemize}
\item
  \begin{quote}
  \protect\hypertarget{_hhvd696t4e1b}{}{}Assess whether
  event-granularity makes sense in object stores that tend to require
  large chunks of data for efficiency
  \end{quote}
\item
  \begin{quote}
  \protect\hypertarget{_hut1j5u624ly}{}{}Does it improve recoverability
  from preemption, in particular when using cloud spot resources and/or
  dynamic HPC resources
  \end{quote}
\end{itemize}

@@

\begin{itemize}
\item
  \begin{quote}
  \protect\hypertarget{_2uzvesmlywml}{}{}row-based vs. column-based
  organization: impact of storage organization on the performance of
  each kind of access, potential storage format providing good
  performance for both
  \end{quote}
\item
  \begin{quote}
  \protect\hypertarget{_6xwv25emxnty}{}{}Can map-reduce, Spark-like
  analysis and their functional or declarative interfaces be adapted to
  HEP analysis needs?
  \end{quote}
\item
  \begin{quote}
  \protect\hypertarget{_vfqvjf8nrae}{}{}Evaluate just in time
  decompression schemes and mappings onto hardware architectures
  considering the flow of data from spinning disk to memory and
  application
  \end{quote}
\end{itemize}

@@

\begin{itemize}
\item
  \begin{quote}
  \protect\hypertarget{_z9tvlvkzecc1}{}{}Quantify benefit of caching for
  main use cases (reconstruction, analysis, simulation)
  \end{quote}
\item
  \begin{quote}
  \protect\hypertarget{_vyj6hh4exb5m}{}{}Assess benefit of caching for
  Machine Learning-based applications, in particular for the learning
  phase
  \end{quote}
\item
  \begin{quote}
  \protect\hypertarget{_ob8u0lfv3fum}{}{}What are the benefits of a CDN
  or NDN approach beyond what HEP is doing today with their data
  delivery systems?
  \end{quote}
\end{itemize}

@@

\begin{itemize}
\item
  \begin{quote}
  \protect\hypertarget{_tmwkcbnhjw6g}{}{}Which role for
  opportunistic/tactical storage
  \end{quote}
\item
  \begin{quote}
  \protect\hypertarget{_tpsd2vti9z2o}{}{}Role of archival storage
  solutions
  \end{quote}
\end{itemize}

@@

\begin{itemize}
\item
  \begin{quote}
  \protect\hypertarget{_ykqlttltg8mz}{}{}Impact of concentrating the
  data in fewer, larger locations (``data-lake'' approach)
  \end{quote}
\item
  \begin{quote}
  \protect\hypertarget{_ym4ou9y4kpdu}{}{}Impact of an increased use of
  opportunistic compute resources, further from the data
  \end{quote}
\end{itemize}

>>>>

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item \textbf{Big Data:} The HL-LHC era will significantly increase both the
  date rate and the data volume. The computing systems will need to
  handle this without significant cost increases and within evolving
  storage technology limitations.
\item \textbf{Dynamic Distributed Computing:} In addition, the significantly
  increased computational requirements for the HL-LHC era will also
  place new requirements on data. Specifically, the use of new types of
  compute resources (\emph{e.g.,} cloud, HPC) with different dynamic
  availability and characteristics are used will require more dynamic
  DOMA systems.
\item \textbf{New Applications:} New applications, such as machine learning
  training or high-rate data-query systems for analysis, will likely be
  employed to meet the computational constraints and to extend the
  physics reach of the HL-LHC. These new applications will place new
  requirements on how and where data is accessed and produced. For
  example, specific applications (e.g. training for machine learning)
  may require use of specialized processor resources such as GPUs,
  placing further requirements on data.
\end{enumerate}

@@

\begin{itemize}
\item \textbf{Data Organization:} This is essentially how data is structured
  as it is written. Most data is written in flat files, in ROOT format,
  typically with a column-wise organization of the data. The records
  corresponding to these columns are compressed. The internal details of
  this organization are typically visible only to individual software
  applications.
\item \textbf{Data Management:} The key challenge here was the transition to
  the use of distributed computing in the form of the grid. The
  experiments developed dedicated data transfer and placement systems,
  along with catalogs, to move data between computing centers. To first
  order, the computing models were rather static: data was placed at
  sites and the relevant compute jobs were sent to the right locations.
  Applications might interact with catalogs or, at times, the workflow
  management systems does this on behalf of the applications.
\item \textbf{Data Access:} Various protocols are used for direct reads
  (rfio, dcap, xrootd, etc.) with a given computer center and/or
  explicit local stage-in and caching for read by jobs. Application
  access may use different protocols than those used by the data
  transfers between sites.
\end{itemize}

@@

\item \textbf{Reconstruction:} Event reconstruction is CPU limited, due to

@@

\item \textbf{Simulation:} HEP event simulation strategies focus on

@@

\item \textbf{Analysis:} Analysis level datasets and subsets are typically

@@

\item \textbf{Replication:} Data replication, whether within a site or

@@

@@

\begin{itemize}
\item Quantify impact on performances
  and resource utilization (storage, network) for the main type of
  access patterns (simulation, reconstruction, analysis)
\item Assess impact on catalogs and
  data distribution
\item Assess whether
  event-granularity makes sense in object stores that tend to require
  large chunks of data for efficiency
\item Does it improve recoverability
  from preemption, in particular when using cloud spot resources and/or
  dynamic HPC resources
\end{itemize}

@@

\begin{itemize}
\item Row-based vs. column-based
  organization: impact of storage organization on the performance of
  each kind of access, potential storage format providing good
  performance for both
\item Can map-reduce, Spark-like
  analysis and their functional or declarative interfaces be adapted to
  HEP analysis needs?
\item Evaluate just in time
  decompression schemes and mappings onto hardware architectures
  considering the flow of data from spinning disk to memory and
  application
\end{itemize}

@@

\begin{itemize}
\item Quantify benefit of caching for
  main use cases (reconstruction, analysis, simulation)
\item Assess benefit of caching for
  Machine Learning-based applications, in particular for the learning
  phase
\item What are the benefits of a CDN
  or NDN approach beyond what HEP is doing today with their data
  delivery systems?
\end{itemize}

@@

\begin{itemize}
\item Which role for
  opportunistic/tactical storage
\item Role of archival storage
  solutions
\end{itemize}

@@

\begin{itemize}
\item Impact of concentrating the
  data in fewer, larger locations (``data-lake'' approach)
\item Impact of an increased use of
  opportunistic compute resources, further from the data
\end{itemize}
